---
output: github_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE
)
```

### Learning Objectives

You will learn how to:

- setup a database connection to WoS-KB with R
- use SQL statements in R Markdown
- assign the results of the SQL query to an R object as a data frame

### Prerequisites

#### Required packages and dependencies

Unfortunately, the KB documentation about how to access the WoS-KB is a bit outdated. Here, we follow the advice from Paul Donner (DZHW) and connect to the WoS-KB using the following packages

```r
install.packages(c("RJDBC", "rJava", "DBI"))
```

In order to connect to the database, you'll need a database driver. You'll find it here:

https://github.com/wiho-fdm/KB-RMarkdown/blob/master/inst/jdbc_driver/ojdbc8.jar

Please download the file `ojdbc8.jar` and put it in the folder `inst/jdbc_driver`

*Some users report frustrations with rJava. An alternative approach would be using an ODBC driver. Here's a brief documentation <https://db.rstudio.com/databases/oracle/>*

We will also use dplyr for analysis

```r
install.packages("dplyr")
```

and, of course, the rmarkdown package.

```r
install.packages("rmarkdown")
```

#### VPN tunnel to FIZ Karlsruhe

Start your VPN connection. Without it, no access to the WoS-KB is possible. 

#### Database credentials 

Keep your database credentials safe by storing it in an `.Renviron`  file 

```r
file.edit("~/.Renviron")
```

and add the following two lines.

```
kb_user=your_user_name
kb_pwd=your_password
```

Restart R.

### Connect to WOS-KB

You first need to establish a connection to the database by

- launching a Java session
- defining where the database driver is stored
- telling the location of the database and your loging credentials safely stored in your `.Renviron` file

```{r setup}
require(RJDBC)
require(rJava)
.jinit()
jdbcDriver <-
  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "inst/jdbc_driver/ojdbc8.jar")
jdbcConnection <-
  dbConnect(
    jdbcDriver,
    "jdbc:oracle:thin:@//biblio-p-db01:1521/bibliodb01.fiz.karlsruhe",
    Sys.getenv("kb_user"),
    Sys.getenv("kb_pwd")
  ) 
```

### Query WOS-KB with SQL

You can make use of this connection in a sql chunk via the connection option.

`````markdown
`r ''````{sql connection=jdbcConnection}
SELECT * FROM wos_b_2019.items
LIMIT 10
```
`````

Real-life example 

```{sql connection=jdbcConnection}
SELECT DOI, UT_EID, PUBYEAR, D_REF_CNT
FROM wos_b_2019.items 
WHERE wos_b_2019.items.ut_eid IN ('000389110200022', '000372645900002', '000400754000138')
```

By default, up to 10 rows will be printed.

### Store query results as R object

The SQL results can be also stored in an R dataframe using the knitr chunk option `output.var`. 

````markdown
`r ''````{sql, connection=jdbcConnection, output.var="my_kb_df"}
SELECT * FROM wos_b_2019.items
LIMIT 10
```
````

```{sql, connection=jdbcConnection, output.var="yearly_collections"}
select
        wos_b_2019.items.pubyear,
        wos_b_2019.databasecollection.edition_value,
        count(distinct(ut_eid)) as pubs                         
    from
        wos_b_2019.items                                            
    inner join
        wos_b_2019.databasecollection                                                                                          
            on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items
    where
    wos_b_2019.items.pubyear in (
            2014, 2015, 2016, 2017, 2018     
        )                                                  
    group by
        wos_b_2019.items.pubyear,
        wos_b_2019.databasecollection.edition_value
```

Inspect `yearly_collections`

```{r}
library(dplyr)
dplyr::glimpse(yearly_collections)
```

Assigning the results of a SQL query to an R data frame allows to further manipulate and analyse the data in R

```{r}
yearly_collections %>%
  group_by(EDITION_VALUE) %>%
  summarise(n = sum(PUBS),
            ylr_mean = mean(PUBS)
) %>%
  arrange(-n)
```


### Recommended readings

- R Markdown: The Definitive Guide. <https://bookdown.org/yihui/rmarkdown/>



